---
title: "Reverse Engineering Project"
author: "Sherwin Esguerra, Kiersten Hacker"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been fatally shot by police since 2015](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(janitor)
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
fatal_police_shooting_data <- read_csv("data/fatal-police-shootings-data.csv")
fatal_police_shooting_agencies <- read_csv("data/fatal-police-shootings-agencies.csv") 
glimpse(fatal_police_shooting_data)
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] After analyzing the data, we found that the total counts for women and men fatally shot and their repsective population percentage comparisons differed from the original story. 121 women with mental illness were shot compared to a total of 366 women fatally shot, which is about 33.06%. For men, 1592 with mental illness were shot out of 7864 total men, which is about 20.24%. However, we realize these changes may be due to updates to the data following the story's release.

```{r}
# Code to find the total number of women fatally shot since 2015
women_shot_by_police <- fatal_police_shooting_data %>% 
  filter(gender == "female") %>% 
  summarise(
    count=n()
  )

# Display results of code below this codeblock
head(women_shot_by_police)
```

```{r}
# Code to find the total number of women with mental illness who were fatally shot
women_mental_illness <- fatal_police_shooting_data %>% 
  filter(was_mental_illness_related == TRUE & gender == "female") %>% 
  summarise(
    count=n()
  )

head(women_mental_illness)
```

```{r}
women_mental_illness_pct <- women_mental_illness %>%
  mutate(
    pct_mentally_ill = women_mental_illness/women_shot_by_police*100
  )
head(women_mental_illness_pct)
```


```{r}
# Code to find the total number of men shot by police
men_shot_by_police <- fatal_police_shooting_data %>% 
  filter(gender == "male") %>% 
  summarise(
    count=n()
  )
head(men_shot_by_police)
```

```{r}
# Code to find the number of men with mental illness who were shot by police
men_mental_illness <- fatal_police_shooting_data %>% 
  filter(was_mental_illness_related == TRUE & gender == "male") %>% 
  summarise(
    count=n()
  )

head(men_mental_illness)
```


```{r}
# Find the percentage of men with mental illness shot
men_mental_illness_pct <- men_mental_illness %>% 
  mutate(
    pct_mentally_ill = men_mental_illness / men_shot_by_police*100
  )

head(men_mental_illness_pct)
```


### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-