---
title: "Reverse Engineering Project"
author: "Sherwin Esguerra, Kiersten Hacker"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been fatally shot by police since 2015](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(janitor)
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
fatal_police_shooting_data <- read_csv("data/fatal-police-shootings-data.csv")
fatal_police_shooting_agencies <- read_csv("data/fatal-police-shootings-agencies.csv") 
glimpse(fatal_police_shooting_data)
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] After analyzing the data, we found that the total counts for women and men fatally shot and their repsective population percentage comparisons differed from the original story. 121 women with mental illness were shot compared to a total of 366 women fatally shot, which is about 33.06%. For men, 1592 with mental illness were shot out of 7864 total men, which is about 20.24%. However, we realize these changes may be due to updates to the data following the story's release.

```{r}
# Code to find the total number of women fatally shot since 2015
women_shot_by_police <- fatal_police_shooting_data %>% 
  filter(gender == "female") %>% 
  summarise(
    count=n()
  )

women_shot_count <- 366

# Display results of code below this codeblock
head(women_shot_by_police)
```

```{r}
# Code to find the total number of women with mental illness who were fatally shot
women_mental_illness <- fatal_police_shooting_data %>% 
  filter(was_mental_illness_related == TRUE & gender == "female") %>% 
  summarise(
    count=n()
  )

head(women_mental_illness)
```

```{r}
women_mental_illness_pct <- women_mental_illness %>%
  mutate(
    pct_mentally_ill = women_mental_illness/women_shot_by_police*100
  )
head(women_mental_illness_pct)
```


```{r}
# Code to find the total number of men shot by police
men_shot_by_police <- fatal_police_shooting_data %>% 
  filter(gender == "male") %>% 
  summarise(
    count=n()
  )
head(men_shot_by_police)
```

```{r}
# Code to find the number of men with mental illness who were shot by police
men_mental_illness <- fatal_police_shooting_data %>% 
  filter(was_mental_illness_related == TRUE & gender == "male") %>% 
  summarise(
    count=n()
  )

head(men_mental_illness)
```


```{r}
# Find the percentage of men with mental illness shot
men_mental_illness_pct <- men_mental_illness %>% 
  mutate(
    pct_mentally_ill = men_mental_illness / men_shot_by_police*100
  )

head(men_mental_illness_pct)
```


### Sentence 2

* **Sentence text**: [Crenshaw said Black women’s deaths also may be dismissed as “collateral damage” if they are killed while police are pursuing someone else. Twenty of the 247 women were killed in that kind of situation, analysis shows. In 12 of those 20 shootings, police said the women killed were caught in crossfire or shot accidentally.]
* **Analysis summary**: [We were able to confirm that 20 women were killed caught in the crossfire or shot accidentally using a filter that accounts for women who were unarmed as well as shot accidentally. However, it is hard to confirm the second statistic, 12 out of 2020, since it is based on what police say about these incidents.]

```{r}
# Put code to reverse engineer sentence here
women_shot <- fatal_police_shooting_data %>% 
  filter(gender == "female")

# Display results of code below this codeblock
women_shot
```
```{r}
# filter for women shot who were unarmed and accidentally shot
women_crossfire <- women_shot %>% 
  filter(armed_with == "unarmed" & threat_type == "accident") %>% 
  summarise(
    count_crossfire = n()
  )

women_crossfire
```

### Sentence 3

* **Sentence text**: [Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. While about twice as many White women were shot and killed as Black women, White women account for five times the Black female population.]
* **Analysis summary**: [The total number of unarmed women killed by police has risen to 33 since the publication of the story. However, using U.S. census data about female populations, we were able to validate the fact that White women account for five times the Black female population.]

```{r}
# Put code to reverse engineer sentence here
women_shot_by_race <- women_shot %>% 
  filter(armed_with == "unarmed") %>% 
  group_by(race) %>% 
  summarise(
    count_race = n()
  )
# Display results of code below this codeblock
women_shot_by_race
```
```{r}
#find the number of white women and black women in the population

#female population based on total of 334485017 and the percentage of female of 50.5% according to US census data
female_population <- 334485017*0.505

#find the total population of black women based on the percentages
black_female_pop <-female_population*0.136

#find the total population of white women based on the percentages
white_female_pop <- female_population*0.758 

#divide the number of white women by the number of black women to check if white make up about 5 times the black female population
white_women_comp_black_women <- white_female_pop/black_female_pop

```

### Sentence 4

* **Sentence text**: [Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths.]
* **Analysis summary**: [We were able to confirm that black women account for less than 1 percent of overall cases where race was known. We were also able to verify that black women are about 13.6% of the female population. However, they only account for about 18% of women shot and killed. Additionally, black women only 14% of unarmed deaths, which is about half of what the story stated.

```{r}
# Find all cases where race was known
shootings_race_known <- fatal_police_shooting_data %>% 
  filter(!is.na(race)) %>% 
  group_by(race) %>% 
  summarise(
    count_race = n()
  )
# Display results of code below this codeblock
shootings_race_known
```
```{r}
#Find incidents where black women were shot
black_women_shot <- fatal_police_shooting_data %>% 
  filter(race=="B" & gender == "female") %>% 
  summarise(
    count_black_women_shot = n()
  )
black_women_shout_count<- 69
black_women_shot

```

```{r}
#total for number of cases where race was known
shootings_race_total <- fatal_police_shooting_data %>% 
  filter(!is.na(race)) %>% 
  summarise(
    count_races_known = n()
  )
shootings_race_total

#percentage of black women compared to all cases where race was known
black_women_comp_all_races<-69/7135

#black women compared to female population
black_female_pop_perc<-black_female_pop/female_population

```
```{r}
#black women shot compared to all women shot and killed
blk_women_shot_perc <- black_women_shot_count/women_shot_count

#find the total of all cases where someone shot was unarmed
total_unarmed_shot <- fatal_police_shooting_data %>% 
  filter(armed_with == "unarmed") %>% 
  summarise(
    count_unarmed_shot = n()
  )
unarmed_shot_count <- 466
total_unarmed_shot

#black women shot compared to all unarmed who were shot
blk_women_comp_unarmed<- black_women_shot_count/unarmed_shot_count
```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```
