---
title: "lab_03"
author: "Derek Willis"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2023.csv", guess_max=322138) %>% 
  clean_names()

# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv", guess_max=17765) %>% 
  clean_names()
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: I found 85 records of payments to the Prince George's County School system (both public schools and their board of education).

```{r}
pg_schools <- payments %>% 
  filter((str_detect(vendor_name, "PRINCE") & str_detect(vendor_name, "SCHOOL")) | (str_detect(vendor_name, "PRINCE") & str_detect(vendor_name, "EDUCATION")) | (str_detect(vendor_name, "PG") & str_detect(vendor_name, "SCHOOL")))
  
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed?

A2: The fiscal year with the largest total amount of payments was 2022. Additionally, I noticed that the years listed show a general growth pattern, especially within the past 4 years or so. However, 2023 is currently listed as last, which makes sense given that the year just began.

```{r}
pg_schools %>% 
  group_by(fiscal_year) %>% 
  summarise(
    total_amount = sum(amount)
  ) %>% 
  arrange(desc(total_amount))
```
Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3.Examples of vendor names that are similar that should be combined into one are:
1. A MORTON THOMAS & ASSOC INC and A MORTON THOMAS & ASSOCIATES INC
2. A & I INC, A&I  INC and A&I INC.
3. AMERICAN DESIGN ASSOCIATE and AMERICAN DESIGN ASSOCIATES

I would use code to detect the common elements between each name within the column and mutate them into one standard name using case_when. I would also clean up any zip codes with only 4 digits by adding a 0 at the beginning through a case_when statement that uses str functions so that they all end up with 5 digits. 

```{r}
capital_project_umd <- payments %>% 
  filter(str_detect(agency_name, "CAPITAL PROJECT") & str_detect(agency_name, "MARYLAND")) %>% 
  group_by(vendor_name, vendor_zip) %>% 
  summarise(
    total_amount = sum(amount)
  )

```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: Grants funding stem cell research have decreased significantly after 2009, reaching an all-time low during the pandemic.

```{r}
stem_cell_grants <- md_grants_loans %>% 
  filter(str_detect(description, "Stem Cell") | str_detect(description, "STEM CELL")) %>% 
  group_by(fiscal_year) %>% 
  summarise(
    total_amount = sum(amount)
  ) %>% 
  arrange(desc(fiscal_year))
```
